---
title: "Impact of collateral on loan performance"
author: "Ang Li"
date: "June 7, 2018"
output:
  html_document:
    number_sections: true
    toc: true
    prettydoc::html_pretty:
      theme: cayman
      highlight: github
    fig_width: 8
    fig_height: 5
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r LoadPackages, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(DiagrammeR)
library(knitr)
library(kableExtra)
library(formattable)
library(dbplyr)
library(lubridate)
library(forcats)
library(odbc)
library(readxl)
library(here)
library(xlsx)
load(here::here('PerfCollateral.RData'))
```

# Sample used in this analysis

A total 11341 loans that originated no later than **December 2017** vintage were selected. Loan performance of the first six months was considered in current analysis, and `Bad`/`Default` is defined as **reaching 60+DPD within first 6 months**.  

# Collateral impact on loan performance

## New vs. Used

```{r NewUsed, echo=FALSE, message=FALSE, warning=FALSE}
table(alf_collateral_perf[, c('IsBad', 'VehicleType')]) %>%
  as.tibble() %>%
  rename(Collateral = VehicleType) %>%
  group_by(Collateral) %>%
  summarise(
      Count = sum(n)
    , BadRatio = round(n[IsBad == 'Bad']/Count * 100, 2)
  ) %>%
  kable(format = 'html', escape = FALSE, align = c('c', 'c', 'c')) %>%
  kable_styling("hover", full_width = FALSE) %>%
  column_spec(3, width = "1cm")
```

Analysis is going to be restricted on 11130 **Used** loans in following sections.

<P style="page-break-before: always">

## Vehicle Age

```{r vage, echo=FALSE, message=FALSE, warning=FALSE}
used_vage_tbl %>%
  mutate(AvgDaysDelinq = round(AvgDaysDelinq, 2)) %>%
  kable(format = 'html', escape = FALSE, align = c('c', 'c', 'c', 'c', 'c')) %>%
  kable_styling("hover", full_width = FALSE) %>%
  column_spec(5, width = "1cm")

used_vage_tbl %>%
  ggplot(aes(VehicleAge, BadRatio, fill = desc(BadRatio))) +
  geom_col() + 
  # coord_flip() + # rotate bars 
  scale_x_discrete(limits = vehicle_age_tbl$VehicleAge) +
  labs(
    title = '"Bad Ratio" of each vehicle age group',
    caption = '"Bad" is defined as reaching 60+DPD within first 6 months'
  ) + 
  theme(legend.position="none")
```

<P style="page-break-before: always">

## Mileage

Mileage was divided into 8 buckets, each bucket has same number of loans.   

```{r mileage, echo=FALSE, message=FALSE, warning=FALSE}
used_mileage_tbl %>%
  mutate(AvgDaysDelinq = round(AvgDaysDelinq, 2)) %>%
  kable(format = 'html', escape = FALSE, align = c('c', 'c', 'c', 'c', 'c')) %>%
  kable_styling("hover", full_width = FALSE) %>%
  column_spec(5, width = "1cm")

used_mileage_tbl %>%
  ggplot(aes(mileage_buckets, BadRatio, fill = desc(BadRatio))) +
  geom_col() + 
  labs(
    title = '"Bad Ratio" of each mileage bucket',
    caption = '"Bad" is defined as reaching 60+DPD within first 6 months'
  ) + 
  theme(legend.position="none")
```

<P style="page-break-before: always">

## Vehicle Estimated Value

Vehicle estimated value was divided into 8 buckets, each bucket has same number of loans. 

```{r vvalue, echo=FALSE, message=FALSE, warning=FALSE}
used_value_tbl %>%
  mutate(AvgDaysDelinq = round(AvgDaysDelinq, 2)) %>%
  kable(format = 'html', escape = FALSE, align = c('c', 'c', 'c', 'c', 'c')) %>%
  kable_styling("hover", full_width = FALSE) %>%
  column_spec(5, width = "1cm")

used_value_tbl %>%
  ggplot(aes(value_buckets, BadRatio, fill = desc(BadRatio))) +
  geom_col() + 
  labs(
    title = '"Bad Ratio" of each value bucket',
    caption = '"Bad" is defined as reaching 60+DPD within first 6 months'
  ) + 
  theme(legend.position="none")
```

<P style="page-break-before: always">

## Loan To Value

`LTVPct` was divided into 8 buckets, each bucket has same number of loans.

```{r ltv, echo=FALSE, message=FALSE, warning=FALSE}
used_ltv_tbl %>%
  mutate(AvgDaysDelinq = round(AvgDaysDelinq, 2)) %>%
  kable(format = 'html', escape = FALSE, align = c('c', 'c', 'c', 'c', 'c')) %>%
  kable_styling("hover", full_width = FALSE) %>%
  column_spec(5, width = "1cm")

used_ltv_tbl %>%
  ggplot(aes(ltv_buckets, BadRatio, fill = desc(BadRatio))) +
  geom_col() + 
  labs(
    title = '"Bad Ratio" of each LTV bucket',
    caption = '"Bad" is defined as reaching 60+DPD within first 6 months'
  ) + 
  theme(legend.position="none")
```